\documentclass{beamer}

% Theme
\usetheme{}
% default theme was Madrid

% Optional packages
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath, amsfonts, amssymb}

% Title information
\title[Short Title]{Computation for Foundation Models}
\subtitle{Tensor Paralellism}
\author{authors}
\institute{}
\date{\today}

\begin{document}

% Title Page
\begin{frame}
  \titlepage
\end{frame}


% Section 2
\section{More Content}
\begin{frame}{What is a foundation model}
  A foundational model is a machine/ deep learning model that can be applied
  to a wide range of use cases. Examples include GPT, stable diffusion, and DALL-E.
  The most common architecture used currently is the transformer architecture.

\end{frame}

\begin{frame}{Simplified Foundation Model}
    To begin with, we will only look at what happens to the input of a foundation model,
    and not the process of it backtracking and learning. This version of a "model" has a lot
    stripped away from it for the purpose of explaining, but the methods shown here not only
    can be used for a full model, but are actively being used in existing models. For parallelism
    analysis, it is common to treat each column of $X$ independently.
    $$ y = f_n(f_{n-1}(...f_1(X))) $$
    $$ f_i = \sigma(W_iX + b_i) $$
    Our input $X$ is a 2d matrix, the W's are linear transformations so they
    can be represented as matrices, and $\sigma$ is a nonlinear function, and y
    is the output.
\end{frame}

\begin{frame}{Tensor Parallelism}
  Tensor parallelism is a way of computing our large from the previous slide across
  multiple processors, or gpus, using linear algebra tricks to split up the computations.

\end{frame}


\begin{frame} {Tensor Parallelism (1D)}
  For our 1D case, $X$ being a matrix doesn't actually matter and we can simplify
  how we think about it to just considering the processing of each column of $X$, which we
  will label $x$.
  So for 1D, our model simplifies to $f_i = \sigma(W_ix +b_i)$.

  $$WX = 
  W
\begin{bmatrix}
x_1 & x_2 & \cdots & x_r
\end{bmatrix}
=
\begin{bmatrix}
Wx_1 & Wx_2 & \cdots & Wx_r
\end{bmatrix}
$$
In other words, we will think of $x$ as being a single column vector for our 1D tensor
parallelism cases.
\end{frame}


\begin{frame} {Tensor Parallelism(Row)}
  We can split the weight matrix $W_i$ into groups of rows depending on how many processors we have.
  Each entry of the $k$th result of $W_i x$ is equal to the $k$th row of $W_i$ times $x$,
  so we just have each processor compute $C_j x$. Since each processor is fully responsible for specific
  entries of the result, we can also give the corresponding entries of $b_i$, which we will label $b_{L_j}$
  to each processor and apply our nonlinear function.
  We end up with each processor computing $\sigma(C_j x + b_{L_j})$, and then gathering the various elements from 
  each device into a single vector to be fed into the next function

  $$ W_i =
  \begin{bmatrix}
  C_1 \\
  C_2 \\
  \vdots \\
  C_m
  \end{bmatrix}
  $$
\end{frame}

\begin{frame} {Tensor Parallelism(Column)}
  We can split the weight matrix $W_i$ into groups of columns depending on the number of processors we have.
  Each $A_j$ is stored on a different processor, and along with it we send the entries of $x$
  that correspond to the columns of $W_i$ that are held within each $A_j$ as a column vector, $x_{L_j}$.
  $$ W_i = \begin{bmatrix} A_1 & A_2 & ... & A_3 \end{bmatrix} $$
  Each processor is then computing $A_j x_{L_j}$ which we can think of as
  $$ A_j x_{L_j} = 
  \begin{bmatrix}
  a_{11}(j) & a_{12}(j) & ... & a_{1k}(j) \\
  a_{21}(j) & a_{22}(j) & ... & a_{2k}(j) \\
  \vdots & \vdots & \ddots & \vdots \\
  a_{m1}(j) & a_{m2}(j) & ... & a_{mk}(j)
  \end{bmatrix}
  \begin{bmatrix}
  x_p \\
  x_p+1 \\
  \vdots \\
  x_p+k-1
  \end{bmatrix}
  $$
  The result of each processor is then combined entry by entry to be create the final vector.

\end{frame}



\begin{frame} {2D Tensor Parallelism}
  Suppose we have $WX + B$ and $q^2$ processors. We can split the matrices W and X
  into $q^2$ blocks each and get
  \[
W =
\begin{bmatrix}
W_{11} & W_{12} & \cdots & W_{1q} \\
W_{21} & W_{22} & \cdots & W_{2q} \\
\vdots & \vdots & \ddots & \vdots \\
W_{q1} & W_{q2} & \cdots & W_{qq}
\end{bmatrix},
\qquad
X =
\begin{bmatrix}
X_{11} & X_{12} & \cdots & X_{1q} \\
X_{21} & X_{22} & \cdots & X_{2q} \\
\vdots & \vdots & \ddots & \vdots \\
X_{q1} & X_{q2} & \cdots & X_{qq}
\end{bmatrix}.
\]

\[
Y = WX =
\begin{bmatrix}
\sum\limits_{k=1}^{q} W_{1k} X_{k1} &
\sum\limits_{k=1}^{q} W_{1k} X_{k2} &
\cdots &
\sum\limits_{k=1}^{q} W_{1k} X_{kq}
\\[6pt]
\sum\limits_{k=1}^{q} W_{2k} X_{k1} &
\sum\limits_{k=1}^{q} W_{2k} X_{k2} &
\cdots &
\sum\limits_{k=1}^{q} W_{2k} X_{kq}
\\[6pt]
\vdots & \vdots & \ddots & \vdots
\\[6pt]
\sum\limits_{k=1}^{q} W_{qk} X_{k1} &
\sum\limits_{k=1}^{q} W_{qk} X_{k2} &
\cdots &
\sum\limits_{k=1}^{q} W_{qk} X_{kq}
\end{bmatrix} + B.
\]
\end{frame}

% TODO FIX THIS THIS IS VERY WRONG RIGHT NOW
\begin{frame}{2D Tensor Parallelism pt 2 (SUMMA algorithm)}

  Since we have a grid of $q^2$ processors, and a matrix of $q^2$ blocks, each processor is given $W_{ij}$ and $X_{ij}$
  and computes $\sum\limits_{k=1}^{q} W_{ik}X_{kj}$. To do this, each processor computes $W_{ij} X_{ij}$ and broadcasts
  the result to the other processors in the same row and column which gives each processor exactly the matrices it needs
  to compute the sum.

  Note: There are similar algorithms to compute both $A^TB$ and $AB^T$, which are are both computations that
  can be found in a transformer model, the most common type of foundation model.
\end{frame}

\begin{frame}{Performance Statistics}
  -2d tensor parallelism has a lower communication cost than 1d

\end{frame}

\begin{frame}{Strengths and Weaknesses}
  2D tensor parallelism comes with the downside of requiring a square number of processors to work.
  Also, if the split is too aggressive and the matrix blocks become
  too small which leads to poor GPU utilization.

  On the other hand, 2D tensor parallelism comes with decreased communication volume which means
  increased communication efficiency. Since we are splitting in both
  dimension for 2D, each processor only needs a smaller block
  of $W$ and $X$ which reduces the amount of data exchanged.

\end{frame}

\begin{frame}{Relevancy, Impact, real world uses and active research}
  As AI models are getting bigger and bigger, we need ways to actually handle these massive models
  both on the hardware and software side, as well as ways to parallelize the models in the first place.

  Tensor parallelism is used in many large foundation models, which are being used in many 
  different ways today. From chat bots and image generation, all the way to more professional use cases
  such as medical ones.

  As for active research on computation for foundation models, almost all of the information
  talked about shows up in distributed training systems for models such as
  DeepSpeed and Megatron-LM. Both of which are actively researching new
  parallelism strategies.

\end{frame}

% Ending
\begin{frame}{Thank You}
  Questions?
\end{frame}

\begin{frame}{Sources}

\end{frame}

\end{document}
